---
title: "`r paste(params$date_tag, params$experiment, sep='_')`"
author: "Jess Vera"
output:
  html_document:
    toc: yes
    df_print: paged
    code_folding: show
params:
  date_tag: '2026-08-14'
  experiment: 'reporting_error_investigation'
---

Start date: `r params$date_tag`    
Last updated: `r Sys.time()`

# Background

While pulling the reporting metrics for July to add to the "master" slide deck 
I noticed that the "total (unique) access requests have been approved for the ARK Portal" 
count month-over-month increase was not matching the July new/renewed counts. 

Further examination shows that March, April, May, and Jun appear to be under reporting 
on the monthly new/renewed stats and I guess also somehow the total uique count is also 
off in some way?

# Goal 

perform an analysis here to review all of the metrics for 2026 and see what needs to be 
fixed in `scrape_approved_access_page.R` to resolve and prevent from happening again.

```
for commit in $(git rev-list HEAD -- catalog_results/AMP\ AIM\ Team\ Permissions.csv); do 
  git show $commit:catalog_results/AMP\ AIM\ Team\ Permissions.csv >     2026-06-25_permissions_audit/team_permissions_history/AMP-AIM_team_permissions.${commit}.csv; 
done
```

# Functions

```{r, functions}

```

# Setup

```{r, setup}
# source rmarkdown_setup.R
source("~/Setup/rmarkdown_setup.R")
Sys.setenv(RETICULATE_PYTHON = "/Users/jvera/micromamba/envs/synapseclient_v4.12.0/bin/python")
reticulate::use_condaenv("synapseclient_v4.12.0")

# Load libraries silently
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(glue)
  library(reticulate)
  library(lubridate)

})

make_subdir("all_approved_csvs")

```

```{r, engine='python'}
import sys
import os
from git import Repo
import os.path
import pandas as pd
from datetime import datetime
from zoneinfo import ZoneInfo
import time

```

# Repo commit review

to get the commit hash of all 2026 commits/workflow executions so I can 
pull the data.frame for each.

```{r, engine='python'}
# 1. Provide the path to your local git repository
repo_path = "/Users/jvera/DCC/ARK Portal/approved_access_requests/" 

# 2. Initialize the Repo object
repo = Repo(repo_path)

target_tz = ZoneInfo("America/Chicago")

# 3. Iterate through all commits on the current active branch
# You can pass a specific branch name like repo.iter_commits('main') if desired.
data = {'message': [], 'date': [], "SHA": [], 'files': []}
for commit in repo.iter_commits():
    # Extract the commit summary or the full message
    commit_message = commit.message.strip() 
    
    # Extract the timezone-aware datetime object
    #commit_date = commit.committed_date
    commit_date = datetime.fromtimestamp(commit.committed_date, tz=target_tz)
    
    commit_hash = commit.hexsha
    
    files = []
    for file in commit.stats.files.keys():
      files.append(file)
    
    # only collect data for commits with changes to csv
    if "all_approved_requests.csv" in files:
      # Print the pulled data
      data['message'].append(commit_message)
      data['date'].append(commit_date)
      data['SHA'].append(commit_hash)
      data['files'].append(", ".join(files))

commit_history = pd.DataFrame(data)
# subset to this years commits
commit_history = commit_history[commit_history['date'].dt.year == 2026]

with open("commit_SHA.txt", "w") as file:
  for SHA in list(commit_history['SHA']):
    a = file.write(SHA + "\n")

```

next get a copy of each version of the csv

```{r, engine='bash'}
while read commit; do git show $commit:../../all_approved_requests.csv > all_approved_csvs/all_approved_requests.${commit}.csv; done < commit_SHA.txt
```

# Review

ok, so now I have Dec 2025 through Aug 1 commits of all_approved_requests.csv.

I think as a first pass I'll just run the code essentially as is and see if the same 
log results are returned.

```{r}
approved_access_log <- function(commit_date, commit, debug = FALSE) {
  today_date <- ymd_hms(commit_date)
  last_month <- today_date - months(1)
  last_month_str <- month(last_month, label = TRUE, abbr = FALSE)
  
  # read in report log
  report_log <- scan("test.approved_requests.monthly_report.log", what = "character", sep = "\n", quiet   = TRUE)
  session_log <- c("--------------", 
                   glue("log date: {today_date}"))
  
  fid <- glue("all_approved_csvs/all_approved_requests.{commit}.csv")
  results <- read.csv(fid, check.names = FALSE) %>% unique()
  n <- nrow(unique(select(results, `Project Lead`, Institution)))
  
  # assess renewed DUCs
  renewed_DUC <- group_by(results, `Project Lead`, Institution) %>%
    tidyr::nest()
  test <- unlist(lapply(renewed_DUC$data, nrow))
  idx <- which(test > 1)
  renewed_DUC <- renewed_DUC[idx, ]
  renewed_DUC <- tidyr::unnest(renewed_DUC, cols = c(data))
  
  session_log <- c(session_log, 
                   glue("{n} total (unique) access requests have been approved for the ARK Portal"), 
                   glue("{length(idx)} access requests have been renewed >1 time."))
  
  # measure stats for the last complete month
  n <- filter(results, month_int == month(last_month) & year == year(last_month)) %>% nrow()
  y <- filter(renewed_DUC, month_int == month(last_month) & year == year(last_month)) %>% nrow()
  session_log <- c(session_log, glue("In {last_month_str} {year(last_month)}, there were {n-y} new   access requests approved and {y} that were renewed."))
  # reporting
  if (!debug){
    report_log <- c(session_log, report_log)
    fid <- "approved_requests.monthly_report.log"
    writeLines(report_log, fid, sep = "\n")
  }
  
  out_message <- paste(session_log, collapse = "\n")
  message(out_message)
}
```

```{r}
commit_history <- py$commit_history
commit_history <- arrange(commit_history, date)

purrr::walk2(commit_history$date, commit_history$SHA, function(x,y) {approved_access_log(commit_date = x, commit = y)})
```


# Session Details

```{r results='asis'}
#Print out session info
R.utils::System$getHostname(); 
sessionInfo()
