---
title: "`r paste(params$date_tag, params$experiment, sep='_')`"
author: "Jess Vera"
output:
  html_document:
    toc: yes
    df_print: paged
    code_folding: show
params:
  date_tag: '2026-08-14'
  experiment: 'reporting_error_investigation'
---

Start date: `r params$date_tag`    
Last updated: `r Sys.time()`

# Background

While pulling the reporting metrics for July to add to the "master" slide deck 
I noticed that the "total (unique) access requests have been approved for the ARK Portal" 
count month-over-month increase was not matching the July new/renewed counts. 

Further examination shows that March, April, May, and Jun appear to be under reporting 
on the monthly new/renewed stats and I guess also somehow the total uique count is also 
off in some way?

# Goal 

perform an analysis here to review all of the metrics for 2026 and see what needs to be 
fixed in `scrape_approved_access_page.R` to resolve and prevent from happening again.

```
for commit in $(git rev-list HEAD -- catalog_results/AMP\ AIM\ Team\ Permissions.csv); do 
  git show $commit:catalog_results/AMP\ AIM\ Team\ Permissions.csv >     2026-06-25_permissions_audit/team_permissions_history/AMP-AIM_team_permissions.${commit}.csv; 
done
```

# Functions

```{r, functions}

```

# Setup

```{r, setup}
# source rmarkdown_setup.R
source("~/Setup/rmarkdown_setup.R")
Sys.setenv(RETICULATE_PYTHON = "/Users/jvera/micromamba/envs/synapseclient_v4.12.0/bin/python")
reticulate::use_condaenv("synapseclient_v4.12.0")

# Load libraries silently
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(glue)
  library(reticulate)
  library(lubridate)

})

make_subdir("all_approved_csvs")

```

```{r, engine='python'}
import sys
import os
from git import Repo
import os.path
import pandas as pd
from datetime import datetime
from zoneinfo import ZoneInfo
import time

```

# Repo commit review

to get the commit hash of all 2026 commits/workflow executions so I can 
pull the data.frame for each.

```{r, engine='python'}
# 1. Provide the path to your local git repository
repo_path = "/Users/jvera/DCC/ARK Portal/approved_access_requests/" 

# 2. Initialize the Repo object
repo = Repo(repo_path)

target_tz = ZoneInfo("America/Chicago")

# 3. Iterate through all commits on the current active branch
# You can pass a specific branch name like repo.iter_commits('main') if desired.
data = {'message': [], 'date': [], "SHA": [], 'files': []}
for commit in repo.iter_commits():
    # Extract the commit summary or the full message
    commit_message = commit.message.strip() 
    
    # Extract the timezone-aware datetime object
    #commit_date = commit.committed_date
    commit_date = datetime.fromtimestamp(commit.committed_date, tz=target_tz)
    
    commit_hash = commit.hexsha
    
    files = []
    for file in commit.stats.files.keys():
      files.append(file)
    
    # only collect data for commits with changes to csv
    if "all_approved_requests.csv" in files:
      # Print the pulled data
      data['message'].append(commit_message)
      data['date'].append(commit_date)
      data['SHA'].append(commit_hash)
      data['files'].append(", ".join(files))

commit_history = pd.DataFrame(data)
# subset to this years commits
commit_history = commit_history[commit_history['date'].dt.year == 2026]

with open("commit_SHA.txt", "w") as file:
  for SHA in list(commit_history['SHA']):
    a = file.write(SHA + "\n")

```

next get a copy of each version of the csv

```{r, engine='bash'}
while read commit; do git show $commit:../../all_approved_requests.csv > all_approved_csvs/all_approved_requests.${commit}.csv; done < commit_SHA.txt
```

# Review

ok, so now I have Dec 2025 through Aug 1 commits of all_approved_requests.csv.

I think as a first pass I'll just run the code essentially as is and see if the same 
log results are returned.

```{r}
approved_access_log <- function(commit_date, commit, debug = FALSE) {
  today_date <- ymd_hms(commit_date)
  last_month <- today_date - months(1)
  last_month_str <- month(last_month, label = TRUE, abbr = FALSE)
  
  # read in report log
  report_log <- scan("test.approved_requests.monthly_report.log", what = "character", sep = "\n", quiet   = TRUE)
  session_log <- c("--------------", 
                   glue("log date: {today_date}"))
  
  fid <- glue("all_approved_csvs/all_approved_requests.{commit}.csv")
  results <- read.csv(fid, check.names = FALSE) %>% unique()
  n <- nrow(unique(select(results, `Project Lead`, Institution)))
  
  # assess renewed DUCs
  renewed_DUC <- group_by(results, `Project Lead`, Institution) %>%
    tidyr::nest()
  test <- unlist(lapply(renewed_DUC$data, nrow))
  idx <- which(test > 1)
  renewed_DUC <- renewed_DUC[idx, ]
  renewed_DUC <- tidyr::unnest(renewed_DUC, cols = c(data))
  
  session_log <- c(session_log, 
                   glue("{n} total (unique) access requests have been approved for the ARK Portal"), 
                   glue("{length(idx)} access requests have been renewed >1 time."))
  
  # measure stats for the last complete month
  n <- filter(results, month_int == month(last_month) & year == year(last_month)) %>% nrow()
  y <- filter(renewed_DUC, month_int == month(last_month) & year == year(last_month)) %>% nrow()
  session_log <- c(session_log, glue("In {last_month_str} {year(last_month)}, there were {n-y} new   access requests approved and {y} that were renewed."))
  # reporting
  if (!debug){
    report_log <- c(session_log, report_log)
    fid <- "approved_requests.monthly_report.log"
    writeLines(report_log, fid, sep = "\n")
  }
  
  out_message <- paste(session_log, collapse = "\n")
  message(out_message)
}
```

```{r}
commit_history <- py$commit_history
commit_history <- arrange(commit_history, desc(date))

purrr::walk2(commit_history$date, commit_history$SHA, function(x,y) {approved_access_log(commit_date = x, commit = y)})
```

ok, I got the same results which makes sense. BUt know to sanity check things and see 
why the wrong numbers are being reported

```{r}
# get most up-to-date df
df <- read.csv("../../all_approved_requests.csv")
df$date <- paste(df$month_str, df$year)
test <- group_by(df, date, month_int, year) %>% summarise(N = n())
test <- arrange(test, year, month_int) %>% select(date, N)
my_kable_table(test, cap = "number of approved requests per month")
```

yeah so the log numbers don't match. One good sanity check would be to look specifically 
at say April 2026 data and see why only 19 are getting reported. I'm curious to know 
if renewals or new DUCs are getting miscalculated

```{r}
df <- read.csv("../../all_approved_requests.csv")
df$date <- paste(df$month_str, df$year)
april <- filter(df, date == "Apr 2026")
renewals <- group_by(df, Project.Lead, Institution) %>% tidyr::nest()

test <- filter(renewals, Project.Lead %in% april$Project.Lead)
```

ok, so one thing I'm already seeing is that for George Tsokos there are slight 
variations in the Institution string, one approval using the BIDMC accronym and 
another using the full name "Beth Israel Deaconess Medical Center"

and Matteo Villa using a longer Institution name in one DUC compared to another... yay!

I wont correct for this now but I've added it as a to-do item after this analysis. I'll 
see about adding it as an issue to the repo as well

```{r}
df <- read.csv("../../all_approved_requests.csv")
df$date <- paste(df$month_str, df$year)
april <- filter(df, date == "Apr 2026")
renewals <- group_by(df, Project.Lead, Institution) %>% tidyr::nest()
renewals$has_renewal <- unlist(purrr::map(renewals$data, function(df) {
  nrow(df) > 1
}))

test <- filter(renewals, Project.Lead %in% april$Project.Lead)
test <- filter(test, Institution %in% april$Institution)
table(test$has_renewal)
```

huh, ugh! 

```{r}
today_date <- ymd_hms(commit_history$date[4])
print(today_date)
last_month <- today_date - months(1)
print(last_month)
last_month_str <- month(last_month, label = TRUE, abbr = FALSE)

n <- filter(df, month_int == month(last_month) & year == year(last_month)) %>% nrow()
print(n)

renewed_DUC <- group_by(df, Project.Lead, Institution) %>% tidyr::nest()
test <- unlist(lapply(renewed_DUC$data, nrow))
idx <- which(test > 1)
renewed_DUC <- renewed_DUC[idx, ]
renewed_DUC <- tidyr::unnest(renewed_DUC, cols = c(data))
y <- filter(renewed_DUC, month_int == month(last_month) & year == year(last_month)) %>% nrow()

glue("In {last_month_str} {year(last_month)}, there were {n-y} new access requests approved and {y} that were renewed.")
```

ok, so my code is pull out the correct number for April, that's 'n', so I'm guessing that 
'y' is the issue... Hmm, ok so when I use the latest version of the csv everything is fine?

```{r}
purrr::walk(commit_history$date, function(x,y) {
  approved_access_log(commit_date = x, commit = y)
  }, y = "c4a9a05d40b9167f420883ecbbd28d1c90af2df9")
```

huh, yeah so using the latest version yields results that would match expectations. 
So then data is missing from the monthly commits that is then added in a later scrape???????

```{r}
Aug1_approved <- read.csv("../../all_approved_requests.csv")
Aug1_approved$date <- paste(Aug1_approved$month_str, Aug1_approved$year)
filter(Aug1_approved, year == 2026) %>% pull(date) %>% table()

May1_approved <- read.csv("all_approved_csvs/all_approved_requests.a4089737ab092767179e2a4ae09fa1c5e6077b4e.csv")
May1_approved$date <- paste(May1_approved$month_str, May1_approved$year)
filter(May1_approved, year == 2026) %>% pull(date) %>% table()
```

Ok, so scraping on May 2 doesn't catch all April approvals, lame!!!! I'm gonna ask 
about how long the delay is in Slack, but I may be able to get a sense for this 
from looking at the data. Wait actually it looks like the issue may be that there are 
"repeat" renewals occurring? cuz the unique count of project leads in the Aug data 
is 19! Gah!!!!!

```{r}
Aug1_approved <- filter(Aug1_approved, date == "Apr 2026")
May1_approved <- filter(May1_approved, date == "Apr 2026")

length(unique(Aug1_approved$Project.Lead))
tbl <- table(Aug1_approved$Project.Lead)
dups <- as.character(names(tbl))[tbl > 1]
dups

filter(Aug1_approved, Project.Lead %in% dups) %>% arrange(Project.Lead)

filter(May1_approved, Project.Lead %in% dups)
```

Ok, so it looks as though one or more of the scrapes captured a transient shift in 
the approved date. I say transient b/c what I currently see on the ARK website are only 
the dates that I see captured on May 2. So now I'll iterate through the 2026 commit 
history of the csv to see when this occurred.

```{r}
results <- purrr::map2(commit_history$date, commit_history$SHA, 
                       function(commit_date, commit, x) {
                         fid <- glue("all_approved_csvs/all_approved_requests.{commit}.csv")
                         df <- read.csv(fid)
                         df$commit_date <- commit_date
                         df <- filter(df, Project.Lead %in% x)
                         return(df)
           }, x = dups)

results <- bind_rows(results) %>% filter(year == 2026)
results <- filter(results, Approved_Date %notin% May1_approved$Approved_Date)
results <- arrange(results, Project.Lead, commit_date)

my_kable_table(df = select(results, Project.Lead, Approved_Date, commit_date), 
               cap = "Results of commit history review to id when transient approval date change occurred")
```

Ok so I think the easiest solution here is to collapse by month+year instead of keeping 
the specific day of the month. So I'll apply that to the latest scrape results and recalcuate 
the total number of DUCs and renewals from that

```{r}
Aug1_approved <- read.csv("../../all_approved_requests.csv")
Aug1_approved$date <- paste(Aug1_approved$month_str, Aug1_approved$year)
before <- nrow(Aug1_approved)
Aug1_approved <- select(Aug1_approved, -Approved_Date) %>% unique()
after <- nrow(Aug1_approved)

print(glue("Starting with {before} rows, ending with {after} rows"))


Jul1_approved <- read.csv("all_approved_csvs/all_approved_requests.25d696d69fb16016d1c17700c592a339f5d81ab8.csv")
Jul1_approved$date <- paste(Jul1_approved$month_str, Jul1_approved$year)
before <- nrow(Jul1_approved)
Jul1_approved <- select(Jul1_approved, -Approved_Date) %>% unique()
after <- nrow(Jul1_approved)

print(glue("Starting with {before} rows, ending with {after} rows"))
Jul1_approved$Institution <- str_replace(Jul1_approved$Institution, 
                                         "BIDMC", 
                                         "Beth Israel Deaconess Medical Center")
Jul1_approved$Institution <- str_remove(Jul1_approved$Institution, 
                                         "Division of Rheumatology and Immunology, Department of Internal Medicine, ")
nrow(unique(select(Jul1_approved, Project.Lead, Institution)))
```

This reduces the total number of DUCs by `r before - after`.

now account for some of the minor institution differences

```{r}
Aug1_approved$Institution <- str_replace(Aug1_approved$Institution, 
                                         "BIDMC", 
                                         "Beth Israel Deaconess Medical Center")
Aug1_approved$Institution <- str_remove(Aug1_approved$Institution, 
                                         "Division of Rheumatology and Immunology, Department of Internal Medicine, ")
total_unique <- nrow(unique(select(Aug1_approved, Project.Lead, Institution)))

print(glue("{total_unique} total (unique) access requests have been approved for the ARK Portal"))
# assess renewed DUCs
renewed_DUC <- group_by(Aug1_approved, Project.Lead, Institution) %>%
  tidyr::nest()
test <- unlist(lapply(renewed_DUC$data, nrow))
idx <- which(test > 1)
renewed_DUC <- renewed_DUC[idx, ]
renewed_DUC <- tidyr::unnest(renewed_DUC, cols = c(data))

print(glue("{length(idx)} access requests have been renewed >1 time."))
```

# Recalculate Metric For Log

Ok, so now that I have found a way to clean up the data I will recalcuate and re-log 
the monthly numbers (only expect the total unique and renewals counts to change) by 
looping through the commit history for 2026 with revised code

```{r}
new_metrics_log <- function(commit_date, commit, debug = FALSE) {
  today_date <- ymd_hms(commit_date)
  last_month <- today_date - months(1)
  last_month_str <- month(last_month, label = TRUE, abbr = FALSE)
  filter_date <- paste(month(today_date, label = TRUE, abbr = FALSE), year(today_date))
  
  # read in report log
  report_log <- scan("revised.approved_requests.monthly_report.log", what = "character", sep = "\n", quiet   = TRUE)
  session_log <- c("--------------", 
                   glue("log date: {today_date}"))
  
  fid <- glue("all_approved_csvs/all_approved_requests.{commit}.csv")
  results <- read.csv(fid, check.names = FALSE) %>% unique()
  results$date <- paste(results$month_str, results$year)
  results <- select(results, -Approved_Date) %>% unique()
  results <- filter(results, date != filter_date)
  
  results$Institution <- str_replace(results$Institution, 
                                     "BIDMC", 
                                     "Beth Israel Deaconess Medical Center")
  results$Institution <- str_remove(results$Institution, 
                                    "Division of Rheumatology and Immunology, Department of Internal Medicine, ")
  
  n <- nrow(unique(select(results, `Project Lead`, Institution)))
  
  # assess renewed DUCs
  renewed_DUC <- group_by(results, `Project Lead`, Institution) %>%
    tidyr::nest()
  test <- unlist(lapply(renewed_DUC$data, nrow))
  idx <- which(test > 1)
  renewed_DUC <- renewed_DUC[idx, ]
  renewed_DUC <- tidyr::unnest(renewed_DUC, cols = c(data))
  
  session_log <- c(session_log, 
                   glue("{n} total (unique) access requests have been approved for the ARK Portal"), 
                   glue("{length(idx)} access requests have been renewed >1 time."))
  
  # measure stats for the last complete month
  n <- filter(results, month_int == month(last_month) & year == year(last_month)) %>% nrow()
  y <- filter(renewed_DUC, month_int == month(last_month) & year == year(last_month)) %>% nrow()
  session_log <- c(session_log, glue("In {last_month_str} {year(last_month)}, there were {n-y} new access requests approved and {y} that were renewed."))
  # reporting
  if (!debug){
    report_log <- c(session_log, report_log)
    fid <- "revised.approved_requests.monthly_report.log"
    writeLines(report_log, fid, sep = "\n")
  }
  
  out_message <- paste(session_log, collapse = "\n")
  message(out_message)
}
```

```{r}
commit_history <- py$commit_history
commit_history <- arrange(commit_history, date)

purrr::walk2(commit_history$date, commit_history$SHA, function(x,y) {new_metrics_log(commit_date = x, commit = y, debug = TRUE)})
```

# To-dos

- clean up data to account for minor differences in Institution strings, commit to repo

# Session Details

```{r results='asis'}
#Print out session info
R.utils::System$getHostname(); 
sessionInfo()
