---
title: "`r paste(params$date_tag, params$experiment, sep='_')`"
author: "Jess Vera"
output:
  html_document:
    toc: yes
    df_print: paged
    code_folding: show
params:
  date_tag: '2026-09-01'
  experiment: 'REST_API_comparison'
---

Start date: `r params$date_tag`    
Last updated: `r Sys.time()`

# Background

I have noticed some issues/inconsistencies with the web scraping of <https://arkportal.synapse.org/Data%20Access>. I started some discussion in Slack 
about this <https://sagebionetworks.slack.com/archives/C0AFCTGEQ/p1787068697221139> and 
was reminded that the content of the portal data access page is propagated via 
`${iduReport?accessRestrictionId=9605913}` in the corresponding wiki page.

It was implied that performing a direct API call would be more accurate and "easier" 
than the web scraping.

# Goal 

Perform the August metrics reporting stats using data pulled via the following 
API POST call: <https://rest-docs.synapse.org/rest/POST/accessRequirement/requirementId/approvedSubmissionInfo.html>

# Functions

```{r, functions}

```

# Setup

```{r, setup}
# source rmarkdown_setup.R
source("~/Setup/rmarkdown_setup.R")
Sys.setenv(RETICULATE_PYTHON = "/Users/jvera/micromamba/envs/synapseclient_v4.13.0/bin/python")
reticulate::use_condaenv("synapseclient_v4.13.0")

# Load libraries silently
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(glue)
  library(reticulate)
  library(lubridate)
})

```

```{r, engine='python'}
import sys
import synapseclient
import os.path
import pandas as pd
import synapseutils
import json
from synapseclient.models import Table, SchemaStorageStrategy, query, Column, ColumnType, FacetType, Dataset, Folder, File

sys.version
syn = synapseclient.login()
```

# Main

```{r, engine='python'}
pages = list(range(50, 1000, 50))

results = []

for page in pages:
  page_token = f"50a{page}"
  response_body = "".join(["{", f"accessRequirementId: '9605913', nextPageToken: '{page_token}'", "}"])
  print(response_body)
  response = syn.restPOST(uri = "https://repo-prod.prod.sagebase.org/repo/v1/accessRequirement/9605913/approvedSubmissionInfo", body = response_body)
  results = results + response['results']
  
  if "nextPageToken" not in response.keys():
    break

```

```{r, engine='python'}
data = {'Project Lead' : [], 'Institution': [], 'Datetime': []}
for DUC in results:
  data['Project Lead'].append(DUC['projectLead'])
  data['Institution'].append(DUC['institution'])
  data['Datetime'].append(DUC['modifiedOn'])

df = pd.DataFrame(data)
```

# Results

```{r}
# read in catalog of all previously scraped approved requests
fid <- "../../all_approved_requests.csv"
all_approved <- read.csv(fid, check.names = FALSE)
all_approved$date <- paste(all_approved$month_str, all_approved$year)

today_date <- today()
last_month <- today_date - months(1)
last_month_str <- month(ymd(last_month), label = TRUE, abbr = FALSE)
filter_date <- paste(month(last_month, label = TRUE, abbr = TRUE), year(last_month))

session_log <- c("--------------", 
                 glue("log date: {Sys.time()}"))


# format info of interest into df
results <- py$df
results$Institution <- str_replace(results$Institution, "BIDMC", "Beth Israel Deaconess Medical Center")
results$Institution <- str_remove(results$Institution, 
                                  "Division of Rheumatology and Immunology, Department of Internal Medicine, ")
results <- mutate(results,
                  year = year(ymd_hms(Datetime)),
                  month_str = month(ymd_hms(Datetime), label = TRUE, abbr = TRUE), 
                  month_int = month(ymd_hms(Datetime)))
results$Approved_Date <- unlist(purrr::map(results$Datetime, function(x){
                            out <- paste(c(month(ymd_hms(x)), 
                                           day(ymd_hms(x)), 
                                           year(ymd_hms(x))),
                                           collapse = "/")
                          }))
results$date <- paste(results$month_str, results$year)

results <- bind_rows(all_approved, results) %>% unique()
n <- nrow(unique(select(results, `Project Lead`, Institution)))
session_log <- c(session_log, 
                 glue("{n} total (unique) access requests have been approved for the ARK Portal"))

# assess renewed DUCs
renewed_DUC <- select(results, `Project Lead`, Institution, date) %>% unique() %>%
                  group_by(`Project Lead`, Institution) %>% tidyr::nest()
test <- unlist(lapply(renewed_DUC$data, nrow))
idx <- which(test > 1)
renewed_DUC <- renewed_DUC[idx, ]
renewed_DUC <- tidyr::unnest(renewed_DUC, cols = c(data))

session_log <- c(session_log,
                 glue("{length(idx)} access requests have been renewed >1 time."))

# measure stats for the last complete month
n <- filter(results, date == filter_date) %>% select(`Project Lead`, Institution) %>% unique() %>% nrow()
y <- filter(renewed_DUC, date == filter_date) %>% nrow()
session_log <- c(session_log, glue("In {last_month_str} {year(last_month)}, there were {n-y} new access requests approved and {y} that were renewed."))

session_log
```

# Session Details

```{r results='asis'}
#Print out session info
R.utils::System$getHostname(); 
sessionInfo()
